\documentclass[../main.tex]{subfiles}

\begin{document}
Each block $B^r$ is associated with a \emph{balance table} $S^r$, which captures current information about all accounts in the system. 
$S^0$ is computed from $B^0$;
$S^r$ is computed from $B^r$ and $S^{r-1}$. 

The balance table $S^r$ contains a balance entry for each user in the system.
The balance entry for a user $u$ records the following information:
\begin{itemize}
\item $\APK_u$: account public key (the corresponding secret key is kept privately by the user). 
\item $b_u^r$: account balance.
\item $u$'s status.
\end{itemize}
A user's balance entry may optionally include additional information, which is needed to participate in the consensus protocol:
\begin{itemize}
\item $\VRFPK_u$: Verifiable Random Function (VRF) public key used for cryptographic sortition.
\item $\PartPK_u$: participation signing public key. 
\item $[v,w]$ : A range of rounds for which the user's participation signing key is valid. 
\end{itemize}
$\VRFPK$ and $\PartPK$ are described in more detail in Section~\ref{section:KeyManagement}; 
as usual, their corresponding secret keys are kept privately by the user.

\paragraph{User status}

The status of user $u$ is determined as follows:
\begin{itemize}
\item A user $u$ is considered \emph{online} at round $r$ if and only if she had a nonempty $\VRFPK_u$ specified in her balance entry at round $r-322$ and her $\PartPK_u$ is still valid at round $r$. 
(Here, $322$ refers to a balance lock-back parameter).
\item Otherwise, $u$ is considered \emph{offline} and doesn't participate in round $r$ consensus.
\end{itemize} 

\paragraph{Minimum account balance.}
Balances in the account entries must maintain a minimum account balance of $0.1$ Algo. This restriction is set to mitigate account spamming attacks.

\paragraph{Smallest Algo amount.}
The smallest Algo amount the system keeps is 1 {\em micro Algo}, which is $10^{-6}$ Algo.
\end{document}